---
title: "AguerosR_Assignment1.doc"
author: "Robert Agueros"
date: "5/12/2021"
output: word_document
---

Q1: Load the data set
```{r}
df<-read.csv("https://stats.idre.ucla.edu/stat/data/rdm/patient_pt2_dm.csv"
                  , header=TRUE)
```

Q2: Display the summary of the data
```{r}
summary(df)
```

Q3: Display the first six observations
```{r}
head(df)
```

Q4: Display observations where the patients' ages are greater than 50.
```{r}
sub <- subset(df, df$age>50)
sub
```

Q5: Display only two columns of data: tumorsize and cancerstage
```{r}
dimnames <- df[c("tumorsize", "cancerstage")] 
dimnames
```
Q6: Computer the mean age
```{r}
meanage <- mean(df$age)
meanage
```

Q7: Compute the correlation between tumorsize and age
```{r}
cor(df$tumorsize, df$age)

```

Q8: Plot a histogram for bmi
```{r}
hist(df$bmi)
```

Q9: Plot a scatterplot for age and tumorsize
```{r}
plot(df$age, df$tumorsize)
```

Q10: Display frequency table for ntumors
```{r}
v <- c(df$ntumors)
table(v)
```

