---
title: "AguerosR_Assignment2.doc"
author: "Robert Agueros"
date: "5/20/2021"
output: word_document
---
Q1: Load the data
```{r}
df <- read.csv("https://stats.idre.ucla.edu/stat/data/rdm/patient_pt1_dm.csv")
```

Q2: Display the summary statistics of the data
```{r}
summary(df)
```

Q3: Convert the data type of wbc to numeric.
```{r}
df$wbc <- as.numeric(df$wbc)
is.numeric(df$wbc)
str(df)
```

Q4: Display a frequency table for each of the following categorical variables:
sex, smokinghx, and familyhx. Do you find any invalid values?
Invalid values: df$sex  is 12.2, df$smokinghx is -99, and df$familyhx is -99.
```{r}
table(df$sex)
table(df$smokinghx)
table(df$familyhx)
```

Q5: Display a boxplot for each of the following numerical variables: age, test1, 
test2, co2, and lungcapacity. Do you find any invalid values? 
Yes. Age: 357.89001, Test 1: -99.00000, Test2: -99.00000, Co2: -98, lungcapacity: -99 and -98.
```{r}
boxplot(df$age)
boxplot.stats(df$age)
boxplot(df$test1)
boxplot.stats(df$test1)
boxplot(df$test2)
boxplot.stats(df$test2)
boxplot(df$co2)
boxplot.stats(df$co2)
boxplot(df$lungcapacity)
boxplot.stats(df$lungcapacity)
```

Q6: Recode all invalid values found in previous two steps as missing values (NA). 
```{r}
df$sex[df$sex==12.2] <- NA
df$smokinghx[df$smokinghx== -99] <- NA
df$familyhx[df$familyhx == -99] <- NA
df$age[df$age == 357.8901] <- NA
df$test1[df$test1 == -99.00000] <- NA
df$test2[df$test2 == -99.00000] <- NA
df$co2[df$co2 == -98] <- NA
df$lungcapacity[df$lungcapacity == -99] <- NA
df$lungcapacity[df$lungcapacity == -98] <- NA
```

Q7: Count the number of missing values in the dataset. 
```{r}
sum(is.na(df))
```

Q8: Use its mean to replace missing values for each numeric variable, and use
it most frequent value to replace missing values for each categorical variable. 
```{r}
df$age[is.na(df$age)] <- mean(df$age, na.rm = TRUE)
df$test1[is.na(df$test1)] <- mean(df$test1, na.rm = TRUE)
df$test2[is.na(df$test2)] <- mean(df$test2, na.rm = TRUE)
df$co2[is.na(df$co2)] <- mean(df$co2, na.rm = TRUE)
df$lungcapacity[is.na(df$lungcapacity)] <- mean(df$lungcapacity, na.rm = TRUE)
df$wbc[is.na(df$wbc)] <- mean(df$wbc, na.rm = TRUE)
df$sex[is.na(df$sex)] <- "female"
df$familyhx[is.na(df$familyhx)] <- "no"
df$smokinghx[is.na(df$smokinghx)] <- "never"
```

Q9: Count the number of missing values again to ensures all of them have been replaced. 
```{r}
sum(is.na(df))
```

Q10: Redo the boxplots in step 5 to see what they look like after invalid values have been replaced. 
```{r}
boxplot(df$age)
boxplot(df$test1)
boxplot(df$test2)
boxplot(df$co2)
boxplot(df$lungcapacity)
```

